Supporting Information\\ \textit{Cross Validation Approaches for Penalized Cox Regression}

### 2022-10-27 main.tex
Edits to Section 2

Another approach, more common in the random survival forest literature than the penalized regression literature, is to calculate deviance residuals \citep{Therneau1990}. As noted earlier, if we do not estimate the baseline hazard, then we cannot predict absolute failure time. However, if we combine Cox regression with an estimate of the baseline hazard, such as the Nelson-Aalen estimator \citep{nelson1969, aalen1978}, then we can obtain residuals based on deviance. These residuals can then be used like conventional residuals to estimate cross-validation error. However, we find this approach also to be rather conservative.

@@

@@

Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.  An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood. We refer to this method as the \emph{cross-validated linear predictors} approach. To be more specific, we would first obtain $\hat{\bbeta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:  
\begin{equation}
  \label{eq:cv-lp}
  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\bbeta}^{-k}
\end{equation} 
After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\boldeta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...  \hat{\eta}^{\text{cv}}_{n})$ for the whole sample is obtained. A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
\begin{equation*}
  L(\hat{\boldeta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{\tau \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_\tau)} \right\} ^{\delta_i}.
\end{equation*}
We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\boldeta}^{\text{cv}}).$$

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).  However, if one takes additional steps to estimate the baseline hazard in addition to estimating the coefficients, then such residuals can be calculated.  The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error.   Originally proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.  In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
\begin{align}
  \label{eq:res-m}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
  \label{eq:res-d}
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
\end{align}
where $H_0$ is the cumulative baseline hazard.

In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$.  However, it is not obvious how to obtain a cross-validated version of $H_0$, or to what extent this is necessary.  We experimented with three versions of cross-validated deviance residuals, each with a different choice for $H_0$: (1) The Nelson-Aalen estimator, which does not involve the covariates and is therefore the same across cross-validation folds, (2) the Breslow estimator \citep{Breslow1972}, and (3) the Kalbfleisch-Prentice estimator \citep{Kalbfleisch2011}.  Details for the three approaches are given in the supplementary materials, along with the results of a simulation study comparing them.  Because the Nelson-Aalen baseline hazard performed best in all situations, we use this estimate to construct the cross-validated deviance residuals throughout the rest of the paper.

>>>>

Another approach, more common in the random survival forest literature than the penalized regression literature, is to calculate deviance residuals. As noted earlier, if we do not estimate the baseline hazard, then we cannot predict absolute failure time. However, if we combine Cox regression with an estimate of the baseline hazard, such as the Nelson-Aalen estimator \citep{nelson1969, aalen1978}, then we can obtain residuals based on deviance. These residuals can then be used like conventional residuals to estimate cross-validation error. However, we find this approach also to be rather conservative.

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).  However, if one takes additional steps to estimate the baseline hazard in addition to estimating the regression coefficients, then such residuals can be calculated.  The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error. Originally proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure, particularly in the random survival forest literature \citep{LeBlanc1992, Ishwaran2004}. In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
\begin{align}
  \label{eq:res-m}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
  \label{eq:res-d}
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
\end{align}
where $H_0$ is the cumulative baseline hazard.

In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$. Typically the Nelson-Aalen estimator is used to estimate $H_0$, although other approaches are also possible, such as the Breslow and Kalbfleisch-Prentice estimators \citep{Breslow1972, Kalbfleisch2011}. These are described in greater detail in the supplementary materials, along with the results of a simulation study comparing them. The Nelson-Aalen estimate of the baseline hazard performed best in these simulations, and also seems to be the most common choice in the literature, so we use this estimate to construct cross-validated deviance residuals throughout the rest of the paper.

@@

Both of the cross-validation approaches in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.  An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood. We refer to this method as the \emph{cross-validated linear predictors} approach. To be more specific, we would first obtain $\hat{\bbeta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:  
\begin{equation}
  \label{eq:cv-lp}
  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\bbeta}^{-k}
\end{equation} 
After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\boldeta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...  \hat{\eta}^{\text{cv}}_{n})$ for the whole sample is obtained. A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
\begin{equation*}
  L(\hat{\boldeta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{\tau \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_\tau)} \right\} ^{\delta_i}.
\end{equation*}
We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\boldeta}^{\text{cv}}).$$

@@
